Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-14 shanghai.tex
Describe part 2 in introduction

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
What is a quasicharacter sheaf?
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$, write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a rank-$1$ local system on $\bG$ and $\mu: m^*
\gcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gcs{L} \to \gcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

In Part 1 of this paper we study the category $\QC(G)$ of quasicharacter sheaves.
Since it is a rigid symmetric monoidal category,
and the set of isomorphism classes $\QCiso{G}$ forms a group.
Motivated by the sheaf-function dictionary for commutative, connected
algebraic groups \cite{SGA4.5}, we explore the trace of Frobenius morphism
\[
\QCiso{G} \to \Hom(G(\Fq), \EEx)
\]
and show that it is an isomorphism.

>>>>

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
What is a quasicharacter sheaf?
Let $G$ be a smooth commutative group scheme locally of finite
type over a finite field $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$, write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a rank-$1$ local system on $\bG$ and $\mu: m^*
\gcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gcs{L} \to \gcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

In Part 1 of this paper we study the category $\QC(G)$ of quasicharacter sheaves.
Since it is a rigid symmetric monoidal category,
and the set of isomorphism classes $\QCiso{G}$ forms a group.
Motivated by the sheaf-function dictionary for commutative, connected
algebraic groups \cite{SGA4.5}, we explore the trace of Frobenius morphism
\[
\QCiso{G} \to \Hom(G(\Fq), \EEx)
\]
and show that it is an isomorphism in Theorem \ref{thm:snake}, assuming only
that the component group of $G$ is finitely generated.

We apply Theorem \ref{thm:snake} to the geometrization of quasicharacters
of non-Archimedian tori in part 2.  Let $K$ be a non-Archimedian local field
with finite residue field $\Fq$; we place no restriction on the characteristic of $K$.
Write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$.
Let $T$ be a maximal torus defined over $K$.  The N\'eron model $\TT$ of $T$
is a smooth group scheme over $\OK$ with generic fiber $T$ and a canonical isomorphism
$\TT(\OK) \cong T(K)$.  For each positive integer $d$, let $\GN{T}_d$ be the
Greenberg transform of $\TT \times_{\Spec{\OK}} \Spec{\OK/\pK^d}$; the commutative
group scheme $\GN{T}_d$ plays the r\^{o}le of $G$, above.  Theorem \ref{thm:application}
now provides a geometrization of quasicharacters of $T$ in the form of a canonical isomorphism
\[
\QCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between quasicharacters on $\GN{T}_d$ and characters of $T(K)$ with depth less than $d$.
